### 2013-06-24 trunk/network-science/paper.tex

\begin{abstract}
We address the dynamic properties of typical social networks, in which people
are considered as vertices while relations are considered as edges. We suppose
the evolution of a social network overtime only includes bidirectional new
connections between existing or newly created nodes, and also assume the graph
will evolve as a typical social network. We study the activity and centrality
metrics for nodes in such graphs by regulating a series of reasonable
properties the metrics should satisfy. We then formally define our metrics and
give their mathematical analysis. We have developed a practical based framework
to maintain the evolving large-scale network and the algorithm to compute our
metrics in preferable time and space consumption, including historical queries.
Finally, we use a simple model to generate data and apply our results to such
experiments.\\[1.5em]

\noindent \textbf{Keywords:} social network, dynamic graph, vertex activity,
vertex centrality, online algorithm.
\end{abstract}
% \tableofcontents
\newpage

@@

\begin{theorem}
Evolving activity satisfies integrity.
\end{theorem}
\begin{proof}
	%TODO
\end{proof}

\begin{definition}[Evolving Centrality]
\textbf{Evolving centrality} $\textbf{x}(G_t)$ is a vector in space
$\mathbb{R}^{|V|}$, where
\begin{equation}
\mathbf{x}=\alpha DA^\bullet \mathbf{x} + \beta\mathbf{1},
\end{equation}
$A^\bullet$ is the 0/1 adjacency matrix of $E^\bullet_t$ as defined in section 2,
because centrality accentuate more on the structure of the network but not the
intensity of edges. And
$D$ is a diagonal matrix that
\begin{equation}
D_{ij}=\begin{cases}
0&(i\neq j)\\
\mathit{Act}(G_t,v_i)&(i=j)
\end{cases}
\end{equation}
Then $\textbf{x}_i(G_t)$ is the centrality for node $v_i$ in $G_t$, and is
scale-relative to the influence of $v_i$ in the network at time $t$.
\end{definition}

The equivalent form of evolving centrality $x(G_t,v_i)$ is 
\begin{equation}
x_i(G_t,v_i)=\alpha\sum_jA^\bullet_{ij}
\mathit{Act}(G_t,v_j)x_j(G_t,v_j)+\beta.
\end{equation}

@@

\begin{algorithm}[htbp]
\caption{\textsc{Query Centrality}($t$,$\alpha$,$\beta$,
$\mathbf{x}_0$,$m$)}
\label{alg-cen}
\textbf{Time Complexity:} $O(mK|V|)$, $K$ is the sparsity constant.\\
\textbf{Additional Space:} $O(|V|)$
\begin{algorithmic}[1]
\Require $0<\alpha<\lambda_1(A^\bullet(G_t))^{-1}$ and $\beta>0$ and $m>0$. Here
$\lambda_1(A)$ is the maximal possitive eigenvalue of matrix $A$.
\Ensure return the active centrality $\mathbf{x}(G_t)$.
\For{$i\gets 1\dots|V|$}
\State $a[i]\gets \Call{Query Activity}{t,v_i}$;
\State $\mathbf{x}[i]\gets \mathbf{x}_0[i]$;
\EndFor
\For{$\mathit{count}\gets 1\dots m$}
\For{$i\gets 1\dots|V|$}
\State $\mathbf{x'}[i]\gets \beta$;
\EndFor
\For{$i\gets 1\dots|V|$}
\ForAll{$j\in\mathit{edgeListReduced}[i]$}
\State $\mathbf{x'}[i]\gets \mathbf{x'}[i]+\alpha\cdot a[j]\cdot\mathbf{x}[j]$
\EndFor\EndFor
\For{$i\gets 1\dots|V|$}
\State $\mathbf{x}[i]\gets \mathbf{x'}[i]$;
\EndFor
\EndFor
\State\Return{$\mathbf{x}$};
\end{algorithmic}
\end{algorithm}

@@

What algorithm \ref{alg-cen} do is only the multiplication of a sparse matrix
to a vector, which is a classical problem called sparse matrix-vector
multiplication (SpMV). This problem can be easily solve in both
throughput-oriented and latency-oriented processors in
parallel\cite{williams2009optimization, bell2008efficient,
bell2009implementing, catalyurek1999hypergraph, zhuo2005sparse}.  This is a
good news to us because nowadays parallel machine is much more common in
company, research facilities and even with desktop computer, thanks to the
emergence of GPGPU platforms such as CUDA and OpenCL.

\section{Experiment}
\subsection{Data Generation} 
Our data generation is base on real social networks provided by \cite{database1,
database2}. Unfortunately, those databases do not contain evolving history
recorded data, so we simulate the evolving history by use the
authentic graph as the final state of the tested dynamic graph.
We give a random permutation to the edges, in a way such that there will always
be just one component -- known as the giant component. Because the final state
is from real world data, it is ensured 
the evolving graph generated by our model will satisfy all the property that we
discussed in section 2.2. In addition, Kumar introduced another way to generate
artificial data of evolving social networks\cite{kumar2010structure}.

@@

\subsection{Com}

\section{Conclusion and Perspective}
% TODO

>>>>

\begin{abstract}
We address the dynamic properties of typical social networks, in which people
are considered as vertices while relations are considered as edges. We suppose
the evolution of a social network overtime includes new
connections between existing or newly created nodes, and also assume the graph
will evolve as a typical social network. We study the activity and centrality
metrics for nodes in such graphs by regulating a series of reasonable
properties the metrics should satisfy. We then formally define our metrics and
give their mathematical analysis. We have developed a practical based framework
to maintain the evolving large-scale network and the algorithm to compute our
metrics including historical queries in preferable time and space consumption.
Finally, we use a simple model to generate evolving data from existing social
network data and compare our metrics to static metrics. The results could imply
that dynamic centrality is more capable in spotting hot social events.\\[1.5em]

\noindent \textbf{Keywords:} social network, dynamic graph, vertex activity,
centrality, online algorithm.
\end{abstract}
% \tableofcontents
\newpage

@@

\begin{theorem}
Evolving activity satisfies integrity.
\end{theorem}
\begin{proof}
First we consider the base case, where we add edges to the vertex $v$ at time $t_i$ and $z=0$ initially.

Denote the sum of weights of these new edges by $w(t_i)$, so $z$ decreases to $-\eta w(t_i)$ from 0. By the definition of $z$, we write $z$ as the function of $t$, which is $z(t)=t-t_i-\eta w(t_i)$. For simplify we choose $\mu=e$ here, it's similar to prove for other values of $\mu$. Now the integrity is as follows:

\begin{equation}
\begin{split}
\int_{t_i}^{t_i+\eta w(t_i)} a(G_t,v) \mathrm{d}t &= \int_{t_i}^{t_i+\eta
w(t_i)} e^{\frac{1}{t-t_i-\eta w(t_i)}} \mathrm{d}t \\
&= \int_{-\eta w(t_i)}^{0} e^{\frac{1}{t}} \mathrm{d}t.
\end{split}
\end{equation}

Notice that $e^{1/t}<1(t<0)$, so it's easy for this direction
\begin{equation}
\int_{-\eta w(t_i)}^{0} e^{\frac{1}{t}} \mathrm{d}t< \eta w(t_i).
\end{equation}

Define $f(x)=\int_{-x}^{0} e^{1/t} \mathrm{d}t - \frac{x}{10}$, from $f'(x)=e^{-1/x}-\frac{1}{10}$, at least we can see $f(x)$ increases when $x\ge 1$. So $f(x)\ge f(1)\approx 0.05>0$.

Now we have the other direction: 
\begin{equation}
\int_{-\eta w(t_i)}^{0} e^{\frac{1}{t}} \mathrm{d}t>\frac{\eta w(t_i)}{10},
\end{equation}
as long as $\eta w(t_i)\ge 1$.

Next we return to a general case.

We show the theorem holds for each "phase", where we call the period of time when $z<0$ a "phase" and a "phase" starts and ends both with $z=0$. Consider a "phase" starting from $t_i$ and ending at $t_j$, some edges whose sum of weights is $w(t_k)$ are added to $v$ at time $t_k(i\le k\le j)$, so $z$ decreases from $z(t_k)$ to $z(t_k)-\eta w(t_k)$ sharply.

One can verify that
\begin{equation}
\int_{t_i}^{t_j} a(G_t,v) \mathrm{d}t< \eta \sum_{k=i}^j w(t_k),
\end{equation}
easily since we still have $\mu^{1/z(t)}<1$.

As for the other direction, notice that at time $t_{k+1}$, there are two cases. One is that $z(t_{k+1})$ has increased equal or above $z(t_k)$, namely we can find some $t_k^*\in (t_k, t_{k+1})$, where $t_k^*$ is defined to satisfy that $z(t_k)=z(t_k^*)$ for all $k$. So we have
\begin{equation}
\int_{t_k}^{t_{k+1}} \mu^{\frac{1}{z(t)}} \mathrm{d}t \ge \int_{t_k}^{t_k^*}
\mu^{\frac{1}{z(t_k)+t-t_k}} - \mu^{\frac{1}{z(t_{k})}}\mathrm{d}t.
\end{equation}
The other case is that $z(t_{k+1})$ is still less than $z(t_{k})$, in other words, we will find $t_k^*>t_{k+1}$.

Now from
\begin{equation}
z(t)=
\begin{cases}
z(t_k)+t-t_k&  t\in (t_k, t_{k+1})\\
z(t_{k+1})+t-t_{k+1}& t\in (t_{k+1}, t_{k+2})
\end{cases}
\end{equation}

we have:
\begin{equation}
\begin{split}
&\int_{t_k}^{t_{k+2}} \mu^{\frac{1}{z(t)}} \mathrm{d}t\\
\ge{}& \int_{t_k}^{t_{k+1}^*} \mu^{\frac{1}{z(t)}} \mathrm{d}t \\
={}&\int_{t_k}^{t_{k+1}} \mu^{\frac{1}{z(t_k)+t-t_k}} - \mu^{\frac{1}{z(t_{k})}}
\mathrm{d}t+ \int_{t_{k+1}}^{t_{k+1}^*} \mu^{\frac{1}{z(t_{k+1})+t-t_{k+1}}}\\& -
\mu^{\frac{1}{z(t_{k+1})}} \mathrm{d}t + (t_{k+1}^*-t_{k+1})\mu^{\frac{1}{z(t_{k+1})}}\\
\ge{}& \int_{t_k}^{t_{k}^*} \mu^{\frac{1}{z(t_k)+t-t_k}} -
\mu^{\frac{1}{z(t_{k})}}\mathrm{d}t + \int_{t_{k+1}}^{t_{k+1}^*}
\mu^{\frac{1}{z(t_{k+1})+t-t_{k+1}}} - \mu^{\frac{1}{z(t_{k+1})}}\mathrm{d}t,
\end{split}
\end{equation}
by an easy induction, one can split a "phase" into integrals in the form 
\begin{equation}
\int_{t_k}^{t_{k}^*} \mu^{\frac{1}{z(t_k)+t-t_k}} - \mu^{\frac{1}{z(t_{k})}}\mathrm{d}t
\end{equation}
 Since $ \mu^{1/z}$ decreases as $z$ becomes larger, we can reduce it to the base case now.

Above all, the theorem holds.
\end{proof}

\begin{definition}[Evolving Centrality]
\textbf{Evolving centrality} $\textbf{x}(G_t)$ is a vector in space
$\mathbb{R}^{|V|}$, where
\begin{equation}
\mathbf{x}=\alpha DLA^\bullet \mathbf{x} + \beta\mathbf{1},
\end{equation}
$A^\bullet$ is the 0/1 adjacency matrix of $E^\bullet_t$ as defined in section 2,
because centrality accentuate more on the structure of the network but not the
intensity of edges. And
$D,L$ are diagonal matrices such that
\begin{equation}
D_{ij}=\begin{cases}
0&(i\neq j)\\
\mathit{Act}(G_t,v_i)&(i=j)
\end{cases}\qquad
L_{ij}=\begin{cases}
0&(i\neq j)\\
1/k^\bullet(G_t,v_i)&(i=j)
\end{cases}.
\end{equation}
Then $\textbf{x}_i(G_t)$ is the centrality for node $v_i$ in $G_t$, and is
scale-relative to the influence of $v_i$ in the network at time $t$.
\end{definition}

The equivalent form of evolving centrality $x(G_t,v_i)$ is 
\begin{equation}
x_i(G_t,v_i)=\alpha\sum_j\frac{A^\bullet_{ij}
\mathit{Act}(G_t,v_j)x_j(G_t,v_j)}{k^\bullet(G_t,v_j)}+\beta.
\end{equation}

@@

\begin{algorithm}[htbp]
\caption{\textsc{Query Centrality}($t$,$\alpha$,$\beta$,
$\mathbf{x}_0$,$m$)}
\label{alg-cen}
\textbf{Time Complexity:} $O(mK|V|)$, $K$ is the sparsity constant.\\
\textbf{Additional Space:} $O(|V|)$
\begin{algorithmic}[1]
\Require $0<\alpha<\lambda_1(A^\bullet(G_t))^{-1}$ and $\beta>0$ and $m>0$. Here
$\lambda_1(A)$ is the maximal possitive eigenvalue of matrix $A$.
\Ensure return the active centrality $\mathbf{x}(G_t)$.
\For{$i\gets 1\dots|V|$}
\State $a[i]\gets \Call{Query Activity}{t,v_i}$;
\State $d[i]\gets \mathrm{size}(\mathit{edgeListReduced}[i])$;
\State $\mathbf{x}[i]\gets \mathbf{x}_0[i]$;
\EndFor
\For{$\mathit{count}\gets 1\dots m$}
\For{$i\gets 1\dots|V|$}
\State $\mathbf{x'}[i]\gets \beta$;
\EndFor
\For{$i\gets 1\dots|V|$}
\ForAll{$j\in\mathit{edgeListReduced}[i]$}
\State $\mathbf{x'}[i]\gets \mathbf{x'}[i]+\alpha\cdot
a[j]\cdot\mathbf{x}[j]/d[j]$
\EndFor\EndFor
\For{$i\gets 1\dots|V|$}
\State $\mathbf{x}[i]\gets \mathbf{x'}[i]$;
\EndFor
\EndFor
\State\Return{$\mathbf{x}$};
\end{algorithmic}
\end{algorithm}

@@

What the core part of algorithm \ref{alg-cen} does is only the multiplication
of a sparse matrix to a vector, which is a classical problem called sparse
matrix-vector multiplication (SpMV). This problem can be easily solve in both
throughput-oriented and latency-oriented processors in
parallel\cite{williams2009optimization, bell2008efficient,
bell2009implementing, catalyurek1999hypergraph, zhuo2005sparse}.  This is a
good news to us because nowadays parallel machine is much more common in
companies, research facilities and even with desktop computers, thanks to the
emergence of GPGPU platforms such as CUDA and OpenCL.

\section{Experiment}
\subsection{Data Generation} 
Our data generation is base on real social networks provided by \cite{database1,
database2}. Unfortunately, those databases do not contain evolving history
recorded data, so we simulate the evolving history by use the
authentic graph as the final state of the tested dynamic graph.
We give a random permutation to the edges, in a way such that:
\begin{itemize}
\item There will always
be just one component -- known as the giant component. We never add an edge out
of the giant component, so the evolving of the network seems to have increasing
vertices joining to the network.
\item The time delay of two adjacent new edges is inversely proportional to the
size of the giant component. So the average activity of any indivitual node do
not change tremendously.
\end{itemize}
Because the final state
is from real world data, it is ensured 
the evolving graph generated by our model will satisfy all the property that we
discussed in section 2.2. In addition, Kumar introduced how to generate
an artificial evolving social networks\cite{kumar2010structure}.

@@

The time efficiency in terms of $|V|$ proves to have logarithm growth as
expected.

\subsection{Centrality Analysis}
Here are some traces of both the evolving centrality and the classical PageRank
centrality of specific nodes.

We generates evolving network data in the same model discussed above, with
$|V|\leq 6\,000$ and $|E|\leq 30\,000$. All iterations run for enough time,
until the maximum magnitude of change in individual centralities (known as the
$\infty$-norm of $\Delta\mathbf{x}$) is smaller than $10^{-4}$ by absolute value.
We also calculate the static PageRank centrality in parallel, which do not have
the activity metric as coefficients. All of our calculation assumes
$\alpha=0.8$, and $\beta=1$.

The centrality trace graph for some representative node, when $\mu=4096$ and
$\eta=1000$:
\begin{center} \begin{tikzpicture}[y=.4cm, x=.5cm, font=\sffamily]
	\draw (0,0) -- coordinate (x axis mid) (21, 0);
	\draw (0,0) -- coordinate (y axis mid) (0, 11);
	\node[below=0.5cm] at (x axis mid) {\scriptsize Time ($\times10^3$)};
	\node[rotate=90, left=0.8cm] at (0, 10) {\scriptsize Magnitude of
	Centrality};
	\foreach \x in {1,2,...,20}
	\draw (\x,1pt) -- (\x,-3pt) node[anchor=north] {\scriptsize \x};
		\foreach \y in {1,2,...,11}
		\draw (1pt,\y) -- (-3pt,\y) node[anchor=east] {\scriptsize \y};
	\draw[color=red!70!black,line width=1pt] plot file {exp-g3-1};
	\draw[color=blue!70!black] plot file {exp-g3-2};
	\begin{scope}[shift={(2,8)}]
	\draw[color=red!70!black,line width=1pt] (0,0) -- (2,0)
		node[right]{\scriptsize evolving centrality};
	\draw[color=blue!50!black] (0,1) -- (2,1)
		node[right]{\scriptsize static centrality};
	\end{scope}
\end{tikzpicture} \end{center}

\begin{center} \begin{tikzpicture}[y=.4cm, x=.5cm, font=\sffamily]
	\draw (0,0) -- coordinate (x axis mid) (21, 0);
	\draw (0,0) -- coordinate (y axis mid) (0, 11);
	\node[below=0.5cm] at (x axis mid) {\scriptsize Time ($\times10^3$)};
	\node[rotate=90, left=0.8cm] at (0, 10) {\scriptsize Magnitude of
	Centrality};
	\foreach \x in {1,2,...,20}
	\draw (\x,1pt) -- (\x,-3pt) node[anchor=north] {\scriptsize \x};
		\foreach \y in {1,2,...,11}
		\draw (1pt,\y) -- (-3pt,\y) node[anchor=east] {\scriptsize \y};
	\draw[color=red!70!black,line width=1pt] plot file {exp-g9-1};
	\draw[color=blue!70!black] plot file {exp-g9-2};
	\begin{scope}[shift={(2,8)}]
	\draw[color=red!70!black,line width=1pt] (0,0) -- (2,0)
		node[right]{\scriptsize evolving centrality};
	\draw[color=blue!50!black] (0,1) -- (2,1)
		node[right]{\scriptsize static centrality};
	\end{scope}
\end{tikzpicture} \end{center}

The centrality trace graph for some representative node, when $\mu=2^{20}$ and
$\eta=2000$:
\begin{center} \begin{tikzpicture}[y=.4cm, x=.5cm, font=\sffamily]
	\draw (0,0) -- coordinate (x axis mid) (21, 0);
	\draw (0,0) -- coordinate (y axis mid) (0, 11);
	\node[below=0.5cm] at (x axis mid) {\scriptsize Time ($\times10^3$)};
	\node[rotate=90, left=0.8cm] at (0, 10) {\scriptsize Magnitude of
	Centrality};
	\foreach \x in {1,2,...,20}
	\draw (\x,1pt) -- (\x,-3pt) node[anchor=north] {\scriptsize \x};
		\foreach \y in {1,2,...,11}
		\draw (1pt,\y) -- (-3pt,\y) node[anchor=east] {\scriptsize \y};
	\draw[color=red!70!black,line width=1pt] plot file {exp-g8-1};
	\draw[color=blue!70!black] plot file {exp-g8-2};
	\begin{scope}[shift={(2,8)}]
	\draw[color=red!70!black,line width=1pt] (0,0) -- (2,0)
		node[right]{\scriptsize evolving centrality};
	\draw[color=blue!50!black] (0,1) -- (2,1)
		node[right]{\scriptsize static centrality};
	\end{scope}
\end{tikzpicture} \end{center}
\begin{center} \begin{tikzpicture}[y=.4cm, x=.5cm, font=\sffamily]
	\draw (0,0) -- coordinate (x axis mid) (21, 0);
	\draw (0,0) -- coordinate (y axis mid) (0, 11);
	\node[below=0.5cm] at (x axis mid) {\scriptsize Time ($\times10^3$)};
	\node[rotate=90, left=0.8cm] at (0, 10) {\scriptsize Magnitude of
	Centrality};
	\foreach \x in {1,2,...,20}
	\draw (\x,1pt) -- (\x,-3pt) node[anchor=north] {\scriptsize \x};
		\foreach \y in {1,2,...,11}
		\draw (1pt,\y) -- (-3pt,\y) node[anchor=east] {\scriptsize \y};
	\draw[color=red!70!black,line width=1pt] plot file {exp-g11-1};
	\draw[color=blue!70!black] plot file {exp-g11-2};
	\begin{scope}[shift={(2,8)}]
	\draw[color=red!70!black,line width=1pt] (0,0) -- (2,0)
		node[right]{\scriptsize evolving centrality};
	\draw[color=blue!50!black] (0,1) -- (2,1)
		node[right]{\scriptsize static centrality};
	\end{scope}
\end{tikzpicture} \end{center}

\section{Conclusion and Perspective}
So far, we have established and verified the framework for analysing an
evolving social network. We defined two metrics for the networks, namely
\textbf{evolving activity} and \textbf{evolving centrality}. The
experiment showed that our model can successful compute both of the metrics in
practical cases.

By the final experiment, we can observe that the increase on magnitude of
evolving centrality is similar to that of the static one. However, when the
influence of a node stops increase so fast, our evolving centrality quickly
pulls back, while the static centrality usually keeps a high value for long
time. In many fields, the social concern are supposed to be sensitive to time
, innovation and activeness, instead of imposing more attention to historical
issues that are once very popular. Those fields such as web-page search
engine[?], ranking of movie actors[?], seeking hot scholar topics, etc.
By using evolving centrality, the increment of actions of a node and its
closely related region will be more notable in increment of its ranking of
influence, whereas the static centrality is more likely to trigger cascade on
social preference[?] because people will have less chance to see and rate new
information. The cascade may prevent high quality information from being
popular, and even worse, hasten the monopolization in the commercial aspect.

The evolving centrality is more sensitive to activities of the surroundings, so
it usually changes more dramatically than static centrality which is more
stable. Dynamic metrics provides new methods for network analysis.
